\section{Structure}
In this section we provide an overview over the rest of our work and summarise the content of each remaining chapter.
\paragraph{Preliminary}

In \hyperref[Preliminary]{[Chapter two]} we provide an overview over the topic of secure multiparty computation. We explain the general concept and different important security assumption. Furthermore we explain the database operations that are of importance four our work. 
\paragraph{Framework Description}
\hyperref[Framework Desciption]{[Chapter three]} contains a detailed description of the frameworks we used to perform our benchmarks. Furthermore, a comparative discussion of the frameworks can be found in the chapter, where we especially focus on usability aspects. The chapter concludes with a description of the frameworks that did not make it into our final selection. 
\paragraph{Benchmarking}
In \hyperref[Benchmarking]{[Chapter four]}, we describe the different metrics for which we collected data. In particular, we list the different tools that helped us to do this and provide an overview how they are used. 
\paragraph{Implementation}
In \hyperref[Implementation]{[Chapter five]}, we describe our implantation. First, we describe our use cases and motivate their design.  Then we describe the concrete implementations with the individual frameworks, including commented excerpts from our source code.
\paragraph{Evaluation}
In \hyperref[Evaluation]{[Chapter six]}, we evaluate the data we have collected. We highlight special features and provide a graphical representation of the data. 
Finally we conclude our work with a discussion of our results.
